Predicate Abstraction for Trace Inclusion

### 2017-07-31 symbolic_nfa_universality/symbolic_subset_construction.tex
little fixes in emp univ trans sys

\newpage
\subsection{Universality}
We will snow ketch a basic universality check. 
As in the case of emptiness, it is incomplete due to the infinite data domain, but additionally also due to the need of working with unbounded histories. 
%
Universality of $T$ corresponds to the validity the formula
$$
\Universal \fdef
\bigwedge_{k\geq 0} \forall \vect x[k]:\Language k(\vect x[k])
=
\bigwedge_{k\geq 0} \forall \vect x[k]\,\exists \vect y[k]:\Run k(\vect x[k],\vect y[k]) \land \final_k(x_{k},y_k) 
$$
(the difference from the formula $\Nonempty$ is the leading $\bigwedge$ and the universal quantification over the visible variables). 
%
Let 
$$
\Configurations k(\vect x[k],x,y) \fdef \exists \vect y[k]:\Run k \land x = x_k \land y = y_k  
$$
be a formula which associates a visible trace $\vect x[k]$ of the length $k$ with the resulting values of $x$ and $y$.
Fixing a particular trace $\vect \dval [k]$ results in a so called \emph{configuration}, 
a constraint over $x$ and $y$ with $x$ fixed at the last value $\dval_k$: 
$$
\subst{\Configurations k}{\vect x[k]}{\vect \dval[k]}(x,y)\ \ .
$$
One may hence understand $\Configurations k$ as a function $\lambda \vect x[k].\,\Configurations k$ from traces to configurations. 
%
%
We may rewrite conjunction $\Universal$ as 
$$
\Universal 
\equiv
\bigwedge_{k\geq 0} \forall \vect x[k]\,\exists y:\Configurations k(\vect x[k],x,y) \land \final(x,y) 
$$
That is, every configuration must have $y$ to satisfy $\final$.
Notice that the universality test does not require all but only the smallest (strongest) configurations.

The task is therefore to compute and test all configurations $\subst{\Configurations {k}}{\vect x[k]}{\vect \dval[k]}$. 
To compute a representation of all configurations, we use the fact that configurations at the distance $k+1$ can be obtained as successors of those at the distance $k$ as follows. 
For every visible trace $\vect \dval [k+1]$,
$$
\subst{\Configurations {k+1}}{\vect x[k+1]}{\vect \dval[k+1]} \equiv 
\subst{(\exists{xy}:\subst{\Configurations {k}}{\vect x[k]}{\vect \dval[k]} \land \tau)}{x',y'}{x,y}\land x = \dval_{k+1}\ \ .
$$
The set of all configurations can be hence computed as the least fixpoint of the transition relation that includes the initial configurations.
%
Notice that the lowest configurations at the distance $k+1$ are successors of the lowest configurations at the distance $k+1$.
%
That is, all smallest configurations are found within the distance $k$, i.e.
$$
\bigwedge_{\ell\geq 0} \forall \vect x[\ell]:\Configurations {\ell} \rightarrow \bigvee_{m\leq k} \exists \vect x[m]: \Configurations {m}
$$
%if and only if the distance $k+1$ does not hold any any new configurations, i.e 
if and only if there are no new or smaller configurations at the distance $k+1$, i.e 
$$
\forall \vect x[k+1]:\Configurations {k+1} \rightarrow \bigvee_{\ell\leq k} \exists \vect x[\ell]: \Configurations {\ell}\ \ .
$$

>>>>

\newpage
\subsection{Universality}
We will now sketch a basic universality check. 
As in the case of emptiness, it is incomplete due to the infinite data domain, but additionally also due to the need of working with unbounded histories. 
%
Universality of $T$ corresponds to the validity the formula
$$
\Universal \fdef
\bigwedge_{k\geq 0} \forall \vect x[k]:\Language k(\vect x[k])
=
\bigwedge_{k\geq 0} \forall \vect x[k]\,\exists \vect y[k]:\Run k(\vect x[k],\vect y[k]) \land \final_k(x_{k},y_k) 
$$
(the difference from the formula $\Nonempty$ is the leading $\bigwedge$ and the universal quantification over the visible variables). 
%
Let 
$$
\Configurations k(\vect x[k],x,y) \fdef \exists \vect y[k]:\Run k \land x = x_k \land y = y_k  
$$
be a formula which associates a visible trace $\vect x[k]$ of the length $k$ with the resulting values of $x$ and $y$.
Fixing a particular trace $\vect \dval [k]\in\domain^k$ results in a so called \emph{configuration}, 
a constraint over $x$ and $y$ with $x$ fixed at the last value $\dval_k$: 
$$
\subst{\Configurations k}{\vect x[k]}{\vect \dval[k]}(x,y)\ \ .
$$
One may hence understand $\Configurations k$ as a function $\lambda \vect x[k].\,\Configurations k$ from traces to configurations. 
%
%
We may rewrite conjunction $\Universal$ as 
$$
\Universal 
\equiv
\bigwedge_{k\geq 0} \forall \vect x[k]\,\exists y:\Configurations k(\vect x[k],x,y) \land \final(x,y) 
$$
That is, every configuration must posses $y$ to satisfy $\final$.
Notice that the universality test does not require to test all but only the smallest (strongest) configurations.

The task is therefore to compute and test all smallest configurations $\subst{\Configurations {k}}{\vect x[k]}{\vect \dval[k]}$. 
To compute a representation of all smallest configurations, we use the fact that configurations at the distance $k+1$ can be obtained as successors of those at the distance $k$ as follows. 
For every visible trace $\vect \dval [k+1]$,
$$
\subst{\Configurations {k+1}}{\vect x[k+1]}{\vect \dval[k+1]} \equiv 
\subst{(\exists{xy}:\subst{\Configurations {k}}{\vect x[k]}{\vect \dval[k]} \land \tau)}{x',y'}{x,y}\land x = \dval_{k+1}\ \ .
$$
The set of all configurations can be hence computed as the least fixpoint of the transition relation that includes the initial configurations.
%
Notice that the smallest configurations at the distance $k+1$ are successors of the smallest configurations at the distance $k+1$.
%

That is, all smallest configurations are found within the distance $k$, i.e.
$$
\bigwedge_{\ell\geq 0} \forall \vect x[\ell]:\Configurations {\ell} \rightarrow \bigvee_{m\leq k} \exists \vect x[m]: \Configurations {m}
$$
%if and only if the distance $k+1$ does not hold any any new configurations, i.e 
if and only if there are no new or smaller configurations at the distance $k+1$, i.e 
$$
\forall \vect x[k+1]:\Configurations {k+1} \rightarrow \bigvee_{\ell\leq k} \exists \vect x[\ell]: \Configurations {\ell}\ \ .
$$
